---
title: "DATA_607_Project_4"
author: "Md Jalal Uddin"
date: "November 15, 2016"
output: html_document
---

```{r}
#Create function to download a package
#Sources:http://stackoverflow.com/questions/9341635/check-for-installed-packages-before-running-install-packages

packages <- function(x){
  x <- as.character(match.call()[[2]])
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x,repos="http://cran.r-project.org")
    require(x,character.only=TRUE)
  }
}

#install required packages by using the created function. 

packages("rmongodb")
packages("rjson")
packages("mongolite")
packages("RMongo")
packages("RMySQL")
packages("DBI")

```

```{r}
#loading the packages

library(RMySQL)
library(RMongo)
library(RJSONIO)
library(rjson)
library(RCurl)
library(plyr)
library(knitr)
library(mongolite)
library(rmongodb)

```

Now, connecting to the MySQL server using the sqlconnect function. MySQL is already running in the machine.

```{r}
driver <- dbDriver("MySQL")
sqlconnect <- dbConnect(driver, user="root", password = "abcd1234", dbname = "flights", host = "localhost")

```

```{r}
#Checking how many tables are there in MySQL database
dbListTables(sqlconnect)

```

NOw, I will run some queries to see the number of row in flights, airlines, planes, weather, and airports tables from flights database. 

```{r}
query <- "select * from flights;"
flights <- dbGetQuery(sqlconnect, query)
head(flights, 10) #show first 10 rows of flights table

```

```{r}
nrow(flights) #show total number of rows of flights table

```

```{r}
query <- "select * from airlines;"
airlines <- dbGetQuery(sqlconnect, query)
head(airlines, 10) #show first 10 rows of airlines table

```

```{r}
nrow(airlines) #show total number of rows of airlines table

```

```{r}
query <- "select * from planes;"
planes <- dbGetQuery(sqlconnect, query)
head(planes, 10) #show first 10 rows of planes table

```

```{r}
nrow(planes) #show total number of rows of planes table

```

```{r}
query <- "select * from airports;"
airports <- dbGetQuery(sqlconnect, query)
head(airports, 10) #show first 10 rows of airports table

```

```{r}
nrow(airports)  #show total number of rows of airports table

```


```{r}
query <- "select * from weather;"
weather <- dbGetQuery(sqlconnect, query)
head(weather, 10) #show first 10 rows of weather table

```

```{r}
nrow(weather) #show total number of rows of weather table

```

Now, disconnecting from MySQL server

```{r}
dbDisconnect(sqlconnect)

dbUnloadDriver(driver)

```

Now, by using devtools packages installing rmongodb packages 

```{r}
#Sources: https://www.r-project.org/nosvn/pandoc/rmongodb.html

library(devtools)
install_github(repo = "mongosoup/rmongodb")
library(rmongodb)

```

Now, connect and load to Mongo db by using mongo.create() function. 

```{r}
mongo <- mongo.create()
mongo.is.connected(mongo)


```

```{r}
#Checking how many database is there. 
mongo.get.databases(mongo)

```

We can see there is two data base created in Mongodb-one is "allflights" and the other one is "mflights". I created "allflights" manually by using Command Promt (CMD). I have imported all five tables's csv file from the machine to Mongodb. I have attached the command and the output result in the github and attached that in project 4. And I created  "mflights" by using r command in rmarkdown and saved all of the five files. 
I have showed here the step by step process for both two ways of creating the two database. 

First, creating allflights database manually and renamed the tables using the following command. 

```{r}
mon_airlines <- "allflights.airlines"
mon_airports <- "allflights.airports"
mon_weather <- "allflights.weather"
mon_planes <- "allflights.planes"
mon_flights<-"allflights.flights"

```

Now, Checking the number of rows for all of the five tables.

```{r}
mongo.count(mongo, mon_planes)
mongo.count(mongo, mon_airports)
mongo.count(mongo, mon_airlines)
mongo.count(mongo, mon_weather)
mongo.count(mongo, mon_flights)

```

Now, comparing the the total number of rows of the each of the five tables in Mongodb with total number of rows of each of the five tables in MySQL server. 

```{r}

mongo.count(mongo, mon_planes) == nrow(planes)
mongo.count(mongo, mon_airports) == nrow(airports)
mongo.count(mongo, mon_airlines)  == nrow(airlines)
mongo.count(mongo, mon_flights)  == nrow(flights)
mongo.count(mongo, mon_weather)  == nrow(weather)

#We can see that all of them are true. 

```

Now, I will export the five tables into Mongodb by using following r command. 
Convert all four tables from flights database to JSON then to Mongodb BSON files. And the other flights table I exported manually since it takes too much time to export. 

```{r}
b_airports <- lapply(split(airports, 1:nrow(airports)), function(x) mongo.bson.from.JSON(toJSON(x)))

b_weather <- lapply(split(weather, 1:nrow(weather)), function(x) mongo.bson.from.JSON(toJSON(x)))

b_airlines <- lapply(split(airlines, 1:nrow(airlines)), function(x) mongo.bson.from.JSON(toJSON(x)))

b_planes <- lapply(split(planes, 1:nrow(planes)), function(x) mongo.bson.from.JSON(toJSON(x)))

```


```{r}
# renamed the files into r markdown

m_airlines <- "mflights.airlines"
m_airports <- "mflights.airports"
m_weather <- "mflights.weather"
m_planes <- "mflights.planes"
m_flights<-"mflights.flights"

```

```{r}
#Loading airlines table records from BSON to MongoDB 
if(mongo.count(mongo,m_airlines) != 0){
  mongo.remove(mongo, m_airlines, criteria = mongo.bson.empty()) 
}

```

```{r}
#Checking to see the records were inserted correctly or not
mongo.insert.batch(mongo, m_airlines, b_airlines)

```

```{r}
#Loading airports table records from BSON to MongoDB 

if(mongo.count(mongo,m_airports) != 0){
  mongo.remove(mongo, m_airports, criteria = mongo.bson.empty()) 
}

```

```{r}
#Checking to see the records were inserted correctly or not
mongo.insert.batch(mongo, m_airports, b_airports)

```

```{r}
#Loading weather table records from BSON to MongoDB 

if(mongo.count(mongo,m_weather) != 0){
  mongo.remove(mongo, m_weather, criteria = mongo.bson.empty()) 
}

```

```{r}
#Checking to see the records were inserted correctly or not
mongo.insert.batch(mongo, m_weather, b_weather)

```

```{r}
#Loading planes table records from BSON to MongoDB 

if(mongo.count(mongo,m_planes) != 0){
  mongo.remove(mongo, m_planes, criteria = mongo.bson.empty()) 
}

```

```{r}
#Checking to see the records were inserted correctly or not
mongo.insert.batch(mongo, m_planes, b_planes)

```

Couting the row number for all the five tables exported into Mongodb

```{r}
mongo.count(mongo, m_planes)
mongo.count(mongo, m_airports)
mongo.count(mongo, m_airlines)
mongo.count(mongo, m_weather)
mongo.count(mongo, m_flights)

```

Now, comparing the the total number of rows of the each of the five tables in Mongodb with total number of rows of each of the five tables in MySQL server.

```{r}

mongo.count(mongo, m_planes) == nrow(planes)
mongo.count(mongo, m_airports) == nrow(airports)
mongo.count(mongo, m_airlines)  == nrow(airlines)
mongo.count(mongo, m_weather)  == nrow(weather)
mongo.count(mongo, m_flights)  == nrow(flights)

```

Finally, we I would say the row count have been matched for both mongodb and MySQL data. 

Difference between SQL and NoSQL:

1.SQL databases are relational databases such as Microsoft SQL or MySQl while NoSQL databases are non relational or graphical or distributed datases.

2.SQL databases are table based databases whereas NoSQL databases are document based databases.  
3.SQL databases have predefined schema whereas NoSQL databases have dynamic schema for unstructured data.

4. SQL databases uses SQL ( structured query language ) for defining and manipulating the data, which is very powerful. In NoSQL database, queries are focused on collection of documents.

5. SQL databases are the preferred databases for applications with complex queries while NoSQL databases are the preferred databases for big data.

6. SQL databases are better for applications with heavy transaction processing because it provides better data integrity than NoSQL databases.

Sources: http://www.thegeekstuff.com/2014/01/sql-vs-nosql-db/?utm_source=tuicool
